---
title: "class16"
author: "Victor Yu"
format: pdf
---


```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```

Q1. What column details the total number of people fully vaccinated?
Answer: Column 10

Q2. What column details the Zip code tabulation area?
Answer: Column 2

Q3. What is the earliest date in this dataset?
Answer: 2021-01-05
```{r}
#Use table to look at all the dates and find th earliest one. It's already in order 

table(vax["as_of_date"])
```


Q4. What is the latest date in this dataset? 

Answer: 2023-03-07

```{r}
#Table is already in order by dates/ We can just observe the last row
tail(vax)
```

```{r}
vax$as_of_date[nrow(vax)]
```

```{r}
#Quick overview of dataset
skimr::skim(vax)

#vax$persons_fully_vaccinated
```

Q5. How many numeric columns are in this dataset? 
Answer: 13 columns

Q6: Note that there are "missing values" in the dataset. How many NA values there in the persons_fully_vaccinated column?
Answer: 16621 NA vales

```{r}
#Use 'is.na' to give a T/F matrix
#table () it to give you the count of each

table(is.na(vax$persons_fully_vaccinated))
```


```{r}
#sum () adds up the nukmber of TRUE. We can store this into n.missing to use it
n.missing <- sum (is.na(vax$persons_fully_vaccinated))
round ((n.missing / nrow(vax))*100, 2)
```
Q7. What percent of persons_fully_vaccinated values are missing (to 2 signifcant figures)? 

8.27

## WORKING WITH DATES

```{r}
library (lubridate)
today()
```


```{r}
#Specify that we are using the year-month-day format
#For funsies. This will give an Error! today() - vax$as_of_date[1] need vax$as_of_date first
vax$as_of_date <- ymd (vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```
```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

```{r}
#Find the most recent date in the data set
today() - ymd (vax$as_of_date[nrow(vax)])
```

Q9. How many days have passed since the last update of the dataset? 

Answers: 3 days (as of 3/10/2023)

Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

Answer: 114 unique dates
```{r}
nrow(table(vax$as_of_date))
```

## Zip Codes
```{r}
#installed zipcodeR package
library("zipcodeR")
```

```{r}
#geocode_zip gives certain zip
geocode_zip('92037')
```
```{r}
#Inputting 2 zip codes with zip_distance gives you the distance between them (IN MILES)
zip_distance('92037', '92109')
```

```{r}
# reverse_zipcode pulls out all the related information tied to the zip code
# we can store this in zip_data
zip_data <- reverse_zipcode(c('92037', "92109"))
```

```{r}
#Method 1: Subset SD county
sd <- vax$county == "San Diego"
sdx <- vax[sd,]
```

```{r}
library(dplyr)
```

```{r}
#Method 2

sd.2 <- filter(vax, county == "San Diego")
nrow(sd.2)
```

```{r}
#Keep in mind: sd2 & sdx both "San Diego-sorted" dataframes
nrow(table(sdx$zip_code_tabulation_area))
```

Q11. How many distinct zip codes are listed for San Diego County?
Answer: 107 zip codes in SD

```{r}
#Find index of row largest tot_population
#Use the index to find the zip code is matches with

high <- which.max(sdx$age12_plus_population)
sdx[high, "zip_code_tabulation_area"]
```

Q12. What San Diego County Zip code are ahas the largest 12+ population in this dataset?
Answer:92154 largest population in dataset

```{r}
#Using dplyr to filter the df
sd.date <- filter (vax, county == "San Diego" & as_of_date == "2023-02-28")
```


```{r}
#Remove NA row first & new df for ease

sd.ppfv <- sd.date[is.na(sd.date$percent_of_population_fully_vaccinated) == 0,]

#take average new dataframe without NA rows

mean(sd.ppfv$percent_of_population_fully_vaccinated)
```

Q13. What is the overall average "Percent of Population Fully Vaccinated" value for all San Diego "County" as of "2023-02-28"

Answer: 0.7401687

Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2022-11-15”?

```{r}
#Base R plots
hist(sd.ppfv$percent_of_population_fully_vaccinated, breaks=20)
```

## UCSD & La Jolla

```{r}
#dplyr filter by area code
#T/F dataframe & sdx acutal dataframe that's sorted

ucsd <- filter(sdx, zip_code_tabulation_area == "92037")
ucsd[1,]$age5_plus_population
```

Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area: 

```{r}
library (ggplot2)
```

```{r}
# Fill in the ggplot code from the lab manual
ucsdplot <- ggplot(ucsd) +
 aes(as_of_date,
 percent_of_population_fully_vaccinated) +
 geom_point() +
 geom_line(group=1) +
 ylim(c(0,1)) +
 labs(x="Date", y="Percent Vaccinated")
ucsdplot
```

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2023-02-28")

head(vax.36)
```
Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2023-02-28”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

Anwer: 0.72149

```{r}
mean(vax.36$percent_of_population_fully_vaccinated)
```

```{r}
#Adding existing ucsd.plot with geom_hline()
ucsdplot + geom_hline(yintercept = mean(vax.36$percent_of_population_fully_vaccinated), col = "red", linetype = "dashed")
```

Q17 What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for
ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2023-02-28”?

```{r}
#Fivenum() fucntions works on data., Releases a output of min, lowerQ, median, upperQ, max
fivenum(vax.36$percent_of_population_fully_vaccinated)

#Mean
vax36.mean <- mean(vax.36$percent_of_population_fully_vaccinated)

sixnum <- c(fivenum(vax.36$percent_of_population_fully_vaccinated), vax36.mean)
```

```{r}
summary (vax.36$percent_of_population_fully_vaccinated)
```

Q18. Using ggplot generate a histogram of this data.

```{r}
vax36.plot <- ggplot(vax.36) +
  aes(x=percent_of_population_fully_vaccinated) +
  xlim(0,1) +
  geom_histogram() + geom_vline(aes(xintercept=vax36.mean))
vax36.plot
```
```{r}
#Average for 92040

vax %>% filter(as_of_date == "2023-02-28") %>%
  filter (zip_code_tabulation_area == "92040") %>%
  select(percent_of_population_fully_vaccinated)
```
```{r}
#Average for 92109

vax %>% filter(as_of_date == "2023-02-28") %>%
  filter (zip_code_tabulation_area == "92109") %>%
  select(percent_of_population_fully_vaccinated)
```


Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

Answer:  The two averages are .550469, .69453 which are both below average

Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax36.all <- filter(vax, age5_plus_population > 36144)
ggplot(vax36.all) +
 aes(as_of_date,
 percent_of_population_fully_vaccinated,
 group=zip_code_tabulation_area) +
 geom_line(alpha=0.2, color="blue") +
 ylim(0,1) +
 labs(x="Date", y="Percent Vaccinated",
 title="Vaccination Rates Across CA",
 subtitle="only areas with population above 36k are shown") +
 geom_hline(yintercept = vax36.mean, linetype="dashed")
```

